Demographic Factors and Climate Action: Understanding How Age and Education Shape Sustainable Choices\thanks{Code and data are available at: \url{https://github.com/LexiKnight/toronto_climate/tree/main}.}

### 2024-12-14 paper/climate_change.tex
last minute edits

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Barriers to Green Living: Insights from 2018 and 2021 Toronto Climate Perception Surveys},
  pdfauthor={Lexi Knight},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Barriers to Green Living: Insights from 2018 and 2021 Toronto
Climate Perception Surveys\thanks{Code and data are available at:
\url{https://github.com/LexiKnight/toronto_climate/tree/main}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Uncovering What Stops Sustainable Behaviors and How
Communities Can Overcome Them}
\author{Lexi Knight}
\date{December 3, 2024}

\begin{document}
\maketitle
\begin{abstract}
This study examines the relationship between demographic factors---age,
education, and self-reported knowledge of climate change---and
individuals' likelihood of engaging in climate-friendly behaviors. The
analysis reveals that younger people and those with higher levels of
education are more likely to adopt sustainable actions such as reducing
car use and purchasing electric vehicles. However, the study also
highlights that financial barriers and limited access to resources
remain significant obstacles for broader climate action. These findings
underscore the importance of tailoring policies and interventions to
make sustainable behaviors more accessible to diverse demographic
groups, promoting wider participation in efforts to combat climate
change.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{Introduction}\label{introduction}

Climate change is one of the most urgent challenges we face today. From
rising temperatures to extreme weather events, the planet is already
feeling the consequences of human-driven environmental change
(Intergovernmental Panel on Climate Change (IPCC) 2023). Yet, despite
widespread awareness of these risks, many people still struggle to adopt
behaviors that could help mitigate climate damage (Gifford 2011). This
gap between knowledge and action is a central puzzle that governments,
communities, and organizations are trying to solve (Kollmuss and Agyeman
2002). If we are to tackle this crisis effectively, we need to better
understand what stops people from acting, and how we can motivate change
(Steg 2014).

This paper examines how specific factors---such as age, education, and
self-reported knowledge of climate change---shape people's willingness
to engage in climate-friendly actions (Schultz 2002). By analyzing data
from a 2018 survey (City of Toronto 2024), we explore the connection
between demographic characteristics and behaviors like reducing energy
consumption, using greener transportation, and lowering waste. The
survey also delves into what prevents people from adopting these
behaviors, uncovering obstacles like cost, convenience, and skepticism
about the power of individual actions (Gifford 2011). With this
analysis, we aim to uncover how these factors interact and what they
mean for efforts to encourage sustainable choices (Stern 2000).

The heart of this study is to measure how age, education, and knowledge
of climate change influence the likelihood of adopting certain
behaviors. We found that younger individuals and those with higher
education levels are more likely to engage in actions like reducing car
usage and adopting plant-based diets (Fulton 2022). However, older
individuals are more inclined to take steps like cutting down on
household energy use (Lee 2019). Beyond these demographic trends, we
also discovered that barriers such as financial cost and doubts about
the effectiveness of individual efforts often hold people back from
taking action (Kollmuss and Agyeman 2002). These findings suggest that a
``one-size-fits-all'' approach to climate communication won't work.
Instead, targeted strategies that consider demographic differences are
essential for driving meaningful engagement (Steg 2014).

Understanding these patterns is crucial for crafting effective climate
policies. By identifying which groups are more likely to act and which
face significant barriers, we can better design interventions that speak
to people's concerns and motivations (Stern 2000). The implications of
this study go beyond simply knowing what people think about climate
change---they offer a roadmap for shaping messages and policies that
could shift behavior on a larger scale (Fulton 2022).

The remainder of this paper is structured as follows:
Section~\ref{sec-data} details the data and methods used in our
analysis. Section~\ref{sec-model} describes how we assess the factors
influencing climate-friendly behaviors across demographic groups.
\textbf{?@sec-results} highlights key differences in behavior based on
age, education, and knowledge of climate change.
\textbf{?@sec-discussion} discusses the findings, addresses the study's
limitations, and suggests directions for future research.
Section~\ref{sec-conclusion} concludes with a summary of the key
takeaways and actionable recommendations for enhancing climate action
engagement. Finally, Section~\ref{sec-appendix} contains supplementary
material.

\section{Data}\label{sec-data}

The primary data source for this project is the Climate Perception Study
dataset, provided by the City of Toronto via the Open Data Toronto
platform (City of Toronto 2024). This dataset offers insights into
public perceptions of climate change in Toronto and was accessed from
Open Data Toronto ({``Open Data Toronto''} 2024) on 6 November 2024.
\#\# Software and R-packages This project was conducted using the
statistical software \texttt{R} (R Core Team 2023). Data cleaning and
manipulation were primarily facilitated by the \texttt{tidyverse}
package (Wickham, Averick, et al. 2024), which encompasses various
essential tools. Within this suite, \texttt{dplyr} (Wickham, François,
et al. 2024) was employed for data manipulation tasks such as filtering,
summarizing, and joining datasets, while \texttt{readr} (Wickham,
Hester, et al. 2024) provided efficient functionality for reading and
writing rectangular text data. Specifically, the \texttt{write\_csv}
function from \texttt{tidyverse} was used to save the raw 2018 dataset.
To handle Excel files, we incorporated the \texttt{httr} package for
downloading files directly from online sources and \texttt{readxl}
(Wickham and Bryan 2024), which enabled us to read Excel files
seamlessly. Additionally, the \texttt{openxlsx} package was utilized to
create a new Excel workbook with multiple sheets, allowing us to store
the 2021 dataset in a structured format.

The \texttt{arrow} package was employed to handle large datasets
efficiently, ensuring smooth data processing, while \texttt{forcats} was
used for reordering factor levels to facilitate more insightful analyses
and visualizations. In terms of data tidying, the \texttt{tidyr} and
\texttt{stringr} packages allowed us to reshape and clean the data by
managing character strings and handling missing values. For the final
documentation and report generation, the \texttt{knitr} package was
instrumental in integrating R code and outputs into a cohesive and
reproducible report format. This suite of R packages collectively
enabled us to conduct thorough data management, manipulation, and
analysis, ultimately contributing to the successful completion of the
project.

\subsection{Methodology}\label{methodology}

\subsubsection{Data Collection}\label{data-collection}

The data for this analysis comes from two surveys commissioned by the
City of Toronto to assess residents' perceptions of climate change and
their willingness to take community action. In 2018, Environics Research
conducted n=404 online interviews with Toronto residents aged 18 and
older between October 11 and October 18. Each interview lasted
approximately 10 minutes. The sample was drawn from an online panel,
with quotas set by region, age, and gender to reflect the city's
population based on the 2016 Census, and minor weighting was applied.
Due to the non-random nature of the online survey, no margin of error
was assigned, though a random sample of this size would typically have a
margin of ±4.9 percentage points, 19 times out of 20.

In 2021, a second survey was conducted by Ipsos, a global market
research company. This survey included n=1,400 Toronto residents aged 18
and older and used a mixed-method approach. The sample consisted of
1,000 respondents from an online panel, 300 recruited via cell phone and
landline, and 100 online interviews conducted in Mandarin, Cantonese,
and Punjabi. This combination of methods aimed to enhance the survey's
representativeness. For analysis, Version 2 of the 2021 dataset, which
presents responses in a numeric format, was used.

\subsubsection{Data Analysis}\label{data-analysis}

The primary focus of this analysis was the 2018 dataset, examining six
key variables: age, education, extent informed, likelihood to take
action, reasons for not taking action, and the best method for
delivering information regarding climate change action. I analyzed these
variables at the individual level in the 2018 dataset by calculating the
percentages within each category, which provided a clear understanding
of the distribution and trends within the population.

In addition, I compared the 2018 dataset to the 2021 summary data to
assess changes over the years. By calculating the percentage
distribution for each variable in both years, I was able to identify any
shifts in public awareness, engagement, and preferences regarding
climate change action. This comparison revealed whether there have been
notable changes in factors like the public's level of information about
climate change, their likelihood to take action, and their preferred
communication methods.

This two-step approach allowed for a comprehensive understanding: a
detailed exploration of the 2018 data combined with a broader comparison
of trends between 2018 and 2021.

\subsection{Features}\label{features}

\subsubsection{Age}\label{age}

\paragraph{Individual 2018}\label{individual-2018}

Figure~\ref{fig-one} illustrates the age distribution of survey
respondents in 2018. The histogram displays the frequency of respondents
across different age groups, with the x-axis representing age intervals
and the y-axis indicating the number of respondents within each
interval. The distribution is relatively balanced, with two noticeable
peaks around the 30--40 and 50--60 age ranges, suggesting these age
groups are the most represented in the survey.

The vertical dotted lines in the figure highlight key statistical
measures: the green dotted line represents the median age, while the
blue dashed line indicates the mean age of the respondents. Both
measures fall near the center of the distribution, reinforcing the
observation that the sample is primarily composed of middle-aged
individuals. Although the distribution appears fairly symmetrical, there
is a gradual decline in respondent frequency after the age of 60, with
fewer participants aged 90 and above, indicating a smaller
representation of older individuals.

Overall, Figure~\ref{fig-one} provides valuable context for
understanding the demographic composition of the survey sample. The
predominance of middle-aged respondents may influence the interpretation
of survey results, particularly if age is a factor in attitudes,
behaviors, or preferences related to the study.

@@

\caption{\label{fig-one}The histogram shows the distribution of age
given the 2018 survey respondents.}

@@

\paragraph{Comparison 2018 to 2021}\label{comparison-2018-to-2021}

(\textbf{table-one?}) illustrates the different age ditribution for 2018
and 2021 surveys. The major difference is that there was greater
representation of younger inviduals ranging from 18 to 23 in 2021
compared to 2018. The remaining age groups were not very different.

@@

This table highlights the age distribution of survey respondents in 2018
and 2021.

\subsubsection{Highest Level of Education
Completed}\label{highest-level-of-education-completed}

\paragraph{Individual 2018}\label{individual-2018-1}

Figure~\ref{fig-two} illustrates the distribution of survey respondents
by their highest level of education attained. The most frequently
reported education levels are ``Completed undergraduate degree'' and
``Post graduate/professional school,'' with both categories having a
noticeably higher frequency compared to others, indicating a
well-educated respondent population. In contrast, fewer respondents
reported ``Some community college, vocational, trade school'' and
``Completed community college, vocational, trade school,'' suggesting
these levels are less common among participants. Additionally, a small
portion of respondents preferred not to disclose their education level.
This distribution highlights a skew toward higher educational attainment
within the sample, which may influence responses related to
socio-economic factors and other survey topics.

@@

\paragraph{Comparison 2018 to 2021}\label{comparison-2018-to-2021-1}

(\textbf{table-two?}) illustrates respondant education level in 2018 and
2021 respectively. Please note, the categories of education level are
different for 2018 and 2021 as these were different surveys and
categories were updated in 2021. Despite this, we are still able to make
comparisons. Individuals with high school diplomas or less account for 9
for 2018 and 38 for 2021 (addition of less than highs school and high
school or equivalent categories). This could indicate that the younger
population was included more in the 2021 suvey or that less education
individuals were surveyed. Additionally, its not super clear howver
those with a post graduate or professional degree accounted for 29
percent in 2018 and 35 percent in 2021. This illustrates that more
educated inviduals responded to the 2021 survey.

@@

This table highlights the education distribution of survey respondents
in 2018 and 2021.''

\subsubsection{Extent Feel Informed}\label{extent-feel-informed}

\paragraph{Individual 2018}\label{individual-2018-2}

Figure~\ref{fig-three} depicts the distribution of self-reported climate
change knowledge among survey respondents in 2018. The majority of
participants identified as ``Very informed,'' making this the most
prevalent category by a significant margin. A moderate number of
respondents reported being ``Not very informed,'' while a smaller yet
noticeable group described themselves as ``Extremely informed.'' In
contrast, the categories ``Not at all informed'' and ``Not very
informed'' have the lowest frequencies, with ``Not at all informed''
being particularly rare. This distribution suggests that most
respondents possess a relatively high level of awareness about climate
change, with very few indicating a lack of knowledge on the topic.

@@

\paragraph{Comparison 2018 to 2021}\label{comparison-2018-to-2021-2}

(\textbf{table-three?}) illustrates the self-reported level of extent
inviduals feel informed about climate change and climate action for 2018
and 2021. In 2018, the large marjority of individuals say they are very
informed with 60 percent and the follwoign category is not very informed
with 27 percent. In 2021, The majority of inividuals also say they are
very informed but to a lesser degree with 48 percent and this is closely
followed by only a little informed with 37 percent. There is a trend
that invidiusls in 2018 reported greater levels of knowldege about
climate change whereas 2021 illstrates a more even split between very
informed and only little informed, portraying that fewer individuals
have adequate knowledge about climate change.

@@

This table highlights the extent feel informed distribution of survey
respondents in 2018 and 2021.

\subsubsection{Best Mode of Delivery}\label{best-mode-of-delivery}

\paragraph{Individual 2018}\label{individual-2018-3}

The analysis of the best method for delivering climate change
information portrayed in Figure~\ref{fig-four} revealed that the
majority of individuals in the 2018 dataset preferred receiving
information firstly from advertising campaigns with about 55 percent,
newsletters with about 52 percent, followed by the toronto website wiht
about 46 percent. On the other side of the spectrum, few said they were
uninterested in receiving climate action information with about 5
percent, about 12 percent indicated twitter and then 17 percent
mentioned instagram. Overall social media platforms Facebook, Instagram
and Twtitter were less popular methods of communciation. This could be
because the survey only included individals over the age of 18.
Moreover, the median sample was in the 40s.

@@

\paragraph{Comparison 2018 to 2021}\label{comparison-2018-to-2021-3}

(\textbf{table-four?}) compares responses from 2018 and 2021 surverys
regarding the best method of delivering information on climate change
and climate action. Aside from the additon of four options to the 2021
survey, both surveys questions are the same. The leading category for
2018 is advertising campaigns with 55 percent, closely followed by
newsletter emails with 52 percent and the Toronto website with 46. The
2021 survey illustrates a consistent trend with the same platforms as
the top method of delivering information, specifically the Toronto
website with 44 percent, then enewsletter emails with 43 percent and
advertising campains at 42 percent. These top methods of communciation
are more evenly spread out than the 2018 survey results. Social media
platforms namely twitter, facebook and instagram all increased in the
2021 survey. Notable decreases in term of percentage from 2018 to 2021
include events, enewsletter emails, advertising campaigns, brochures and
pamphlets. The new category of BetterHomesTO.ca webiste accounts for 15
percent of votes for the 2021 survey.

@@

This table highlights the prefered method of communication of survey
respondents in 2018 and 2021.

\subsubsection{Likelihood Taking Action}\label{likelihood-taking-action}

\paragraph{Individual 2018}\label{individual-2018-4}

The likelihood of taking action against climate change was explored by
assessing individuals' intentions and behaviors. Figure~\ref{fig-five}
illustrates self-reported likelihood of taking certain actions that
minimize climate change effects. On the x-axis, we have the nine
different climate actions and on the y-axis is the percentage. The
stacked bar chart allows understanding of how likely individuals are to
take part in the given actions. Sorting waste into the correct bins is
the action that most individuals are doing right now, accounting for 60
percent whith 24 percent saying they are very likely to take part in
this action. The next most prominent actions are walking or cycling
short distances, reduce hydro usage, reduce waste and home improvements.
40 percent of invidiuals report walking or cycling when traveling
shorter distances in the city with 22 percent reporting they are very
likely. However, 7 percent say they are very unlikely to change their
method of transportation. Reducing the amount of energy and water use is
already done by 36 percent and very likely for 29 percent only a mere 3
percent is very unlikely. Reducing personal waste such as contributing
to repurposing used products has similar percentages to responses for
reducing hydro but even fewer individuals that report they are very
unlikely to take action. Similarly, the likelihoods for making home
improvements for energy efficiency such as energy-efficient appliances,
programmable thermostat, LED lightbulbs, green or cool roof. In
contrast, the action where the least number of individuals are
contributing to is the investment of an electric or hybrid vehicle over
gas-powered vehicles with only 8 percent who currently have an energy
efficient vehicle. While 19 percetn are very likely to make this switch,
14 percent say they are very unlikely. This illustrates that buying an
electric or hybrid vehicle is action the least number of individuals are
willing to invest in despite gas-powered vehicles being a major
contributor to greenhouse gases. This action has a huge impact on the
environment and yet 41 percent of this sample say they are unlikely to
change! The second action that individuals are most unlikely to engage
in is eat less meat by incorporating more plant-based foods in their
diet with 12 percent very unlikely and 23 percent somewhat unlikely.
Only 23 percent are already consuming meat alternatives. Vegans
greenhosue gas emissions are on average 37\% lower than those who eat
meat frequently.

@@

\section{Model}\label{sec-model}

The goal of our modelling strategy is twofold. Firstly, we aim to
predict the likelihood of individuals engaging in climate-friendly
actions based on demographic variables, specifically age and education.
The actions considered include investing in electric or hybrid vehicles,
making home improvements (such as using LED lightbulbs), adopting meat
alternatives, minimizing car use, purchasing green products, reducing
hydro usage, reducing waste, sorting waste correctly, and walking or
cycling shorter distances. Secondly, we aim to understand how these
demographic characteristics influence individuals' likelihood of
adopting these behaviors.Here we describe the probability tree model
used to investigate the relationship between likelihood and the
demographic factors of age and education.

@@

In this analysis, we focus on predicting the likelihood of taking
climate-friendly actions based on the demographic variables of age and
education. The likelihood is measured on a five-point Likert scale,
ranging from ``Very unlikely'' to ``Already doing this or have done
this.''

@@

We use a probability tree model to predict the likelihood of individuals
engaging in climate-friendly actions based on two key demographic
variables: age and education.

The general form of the model is:

\textbackslash begin\{align\} y\_i = \beta\_0 + \beta\_1 * Age\_i +
\beta\_2 * Education\_i + \epsilon\_i \textbackslash end\{align*\}
where:

\begin{itemize}
\tightlist
\item
  \(y_i\) represents the likelihood of individual \(i\) engaging in a
  climate-friendly actin, measured on a five point Likert scale:
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  ``Very unlikely''
\item
  ``Somewhat unlikely''
\item
  ``Somewhat likely''
\item
  ``Very likely''
\item
  ``Already doing this or have done this''
\end{enumerate}

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demographic Factors and Climate Action: Understanding How Age and Education Shape Sustainable Choices},
  pdfauthor={Lexi Knight},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Demographic Factors and Climate Action: Understanding How Age and
Education Shape Sustainable Choices\thanks{Code and data are available
at: \url{https://github.com/LexiKnight/toronto_climate/tree/main}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Older Individuals and Those with Higher Education Show
Stronger Commitment to Long-Term Sustainable Actions}
\author{Lexi Knight}
\date{December 13, 2024}

\begin{document}
\maketitle
\begin{abstract}
This study investigates how age and education influence people's
likelihood to adopt climate-friendly behaviors, such as reducing car use
and conserving energy. The analysis shows that younger individuals and
those with higher education are more likely to engage in these actions,
while older people tend to focus on energy-saving behaviors. However,
financial barriers and doubts about the impact of individual actions
prevent many people from taking steps to address climate change. These
findings suggest that targeted policies are needed to make sustainable
choices more accessible to different groups, helping to drive broader
participation in climate action.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\subsection{Introduction}\label{introduction}

Climate change is one of the most pressing issues of our time. Its
effects---rising temperatures, extreme weather events, and ecosystem
disruptions---are already being felt globally, and the urgency for
action has never been higher (Intergovernmental Panel on Climate Change
(IPCC) 2023). Despite widespread awareness of the risks associated with
climate change, many individuals still struggle to adopt behaviors that
could help mitigate its impact. This gap between understanding and
action is a major barrier to effective climate solutions, and
understanding the underlying reasons for this disconnect is essential
for crafting strategies that encourage widespread behavioral change
(Gifford 2011).

This paper seeks to explore how certain demographic
factors---specifically age and education---affect individuals'
likelihood to adopt climate-friendly behaviors. By analyzing data from a
2018 survey (City of Toronto 2024), we aim to understand how these
factors influence behaviors such as reducing energy consumption, using
greener transportation, and minimizing waste. The estimand of this study
is the effect of age and education on the likelihood of adopting these
climate-friendly behaviors. We focus on these variables to understand
whether younger, more educated individuals are more likely to engage in
sustainable actions, and whether targeted strategies could drive greater
participation in climate-positive behaviors.

Through our analysis, we found that younger individuals and those with
higher levels of education were more likely to adopt behaviors such as
reducing car usage and embracing plant-based diets (Fulton 2022).
However, older individuals were more inclined to engage in actions like
cutting down on household energy use (Lee 2019). This suggests that
while age and education are important predictors, the types of
climate-friendly behaviors people adopt may differ across these groups.
Additionally, we identified key barriers to adoption, including
financial cost and doubts about the effectiveness of individual actions,
which often prevent people from taking meaningful steps to reduce their
environmental impact (Kollmuss and Agyeman 2002). These insights
emphasize that a one-size-fits-all approach to climate communication and
action will not be effective; instead, strategies must be tailored to
address the specific needs and motivations of different demographic
groups.

The findings from this study are significant because they provide a
clearer understanding of how demographic characteristics shape the
adoption of climate-friendly behaviors. By identifying which groups are
more likely to act and which face greater barriers, we can design more
effective policies and interventions that speak directly to the concerns
of these populations. This study not only contributes to the broader
understanding of climate change engagement but also offers practical
insights for shaping messages and policies that can inspire individuals
to take action.

The remainder of this paper is structured as follows:
Section~\ref{sec-data} describes the data and methodology used for
analysis, Section~\ref{sec-model} outlines the statistical model used to
assess the effects of age and education on behavior adoption,
Section~\ref{sec-results} presents the key findings, and
Section~\ref{sec-discussion} explores the implications of these results,
addresses limitations, and suggests directions for future research.
Finally, Section~\ref{sec-conclusion} summarizes the main takeaways and
offers recommendations for enhancing climate action engagement.
Additional methodological details and the data cleaning process are
provided in Section~\ref{sec-appendix}.

\subsection{Data}\label{sec-data}

The primary data source for this project is the Climate Perception Study
dataset, provided by the City of Toronto via the Open Data Toronto
platform (City of Toronto 2024). This dataset offers insights into
public perceptions of climate change in Toronto in the years 2018 and
2021. This dataset was accessed from \texttt{Open\ Data\ Toronto}
({``Open Data Toronto''} 2024) on 6 November 2024.

\subsubsection{Software and R-packages}\label{software-and-r-packages}

This project was conducted using the statistical software \texttt{R} (R
Core Team 2023), with several packages facilitating data cleaning,
analysis and reporting. The \texttt{tidyverse} package (Wickham,
Averick, et al. 2024) was central to the project, with \texttt{dplyr}
(Wickham, François, et al. 2024) used in the cleaning and exploratory
data analysis scripts for filtering, summarizing, and joining datasets.
\texttt{readr} (Wickham, Hester, et al. 2024) handled reading and
writing text data, while \texttt{tidyr} (Wickham, Henry, et al. 2024)
reshaped data. \texttt{stringr} (Wickham 2024c) managed character
strings, and \texttt{forcats} (Wickham 2024a) reordered factor levels
for visualizations.

In the download script, \texttt{httr}(Wickham 2024b) retrieved files
from \texttt{Open\ Data\ Toronto} ({``Open Data Toronto''} 2024),
while\texttt{readxl} (Wickham and Bryan 2024) read Excel files, and
\texttt{openxlsx} (Walker et al. 2024) created structured Excel
workbooks. The \texttt{arrow} package (Richardson et al. 2024) optimized
data storage and was used in the simulation and cleaning scripts to save
datasets as Parquet files.

For modeling, \texttt{rpart} (Therneau and Atkinson 2024) and
\texttt{partykit} (Hothorn and Zeileis 2024) built and visualized
decision trees in the model script. The \texttt{testthat} package
(Wickham et al. 2024b) validated simulated and cleaned datasets in their
respective scripts. Visualizations were created using \texttt{ggplot2}
(Wickham et al. 2024a) in the exploratory data analysis. Reporting
relied on \texttt{knitr} (Xie 2024) and \texttt{kableExtra} (Zhu 2024)
to integrate code, outputs, and styled tables. The
\texttt{tinytable}package (Mikolas 2024) generated compact tables in
both the exploratory data analysis script as well as in the paper. These
packages ensured efficient and streamlined data management, analysis and
reporting throughout the project.

\subsubsection{Methodology}\label{methodology}

\paragraph{Data Collection}\label{data-collection}

To embark on this analysis, we turn to two pivotal surveys commissioned
by the City of Toronto to capture residents' perceptions of climate
change and their readiness to take action. Our journey begins in 2018,
when Environics Research conducted an online survey with 404 Toronto
residents aged 18 and older, from October 11 to 18. The sample, drawn
from an online panel, was carefully crafted with quotas based on region,
age, and gender to mirror the 2016 Census. A touch of minor weighting
ensured a more representative reflection of the population. However,
it's important to note that this non-random sample introduces some
potential bias, an aspect we keep in mind as we move forward.

The path continues in 2021 with Ipsos at the helm, stepping in with a
more expansive survey, engaging 1,400 residents from a mix of methods:
1,000 from an online panel, 300 via phone, and 100 through online
interviews conducted in Mandarin, Cantonese, and Punjabi. This diverse
approach sought to enhance the survey's representativeness across
Toronto's varied communities. For our analysis, we used Version 2 of the
2021 dataset, where responses were presented in a clear, numeric format,
offering a solid foundation for our exploration.

With this data in hand, we are now ready to delve into the intricacies
of the findings, unearthing patterns, and understanding the deeper story
they tell about Toronto's residents and their views on climate change
and collective action.

\paragraph{Data Analysis}\label{data-analysis}

The focus of this analysis was the 2018 dataset, which examined five key
variables: age, education, extent informed, likelihood to take action,
and preferred methods for delivering information about climate change
action. These variables were analyzed at the individual level, allowing
for an in-depth exploration of residents' perspectives.

For 2021, summary data was available instead of individual-level data.
To compare trends, I created summary data for 2018. Calculating the
percentage distribution for each variable in both years helped identify
shifts in demographics, public awareness, and engagement, as well as
changes in preferred communication methods. Although the absence of
individual data for 2021 posed a challenge, this comparison still
provided a dynamic view of evolving climate change perspectives in
Toronto. The two-step approach---detailing 2018 data and then comparing
it with 2021 trends---created a comprehensive narrative of change.

\subsubsection{Features}\label{features}

\paragraph{Age}\label{age}

\subparagraph{Individual 2018}\label{individual-2018}

Figure~\ref{fig-one} presents the age distribution of survey respondents
in 2018. The histogram shows the frequency of respondents across age
groups, with age intervals on the x-axis and the number of respondents
on the y-axis. The distribution is fairly balanced, with peaks around
the 30--40 and 50--60 age ranges, indicating that these groups were the
most represented in the survey.

The vertical dotted lines highlight the median and mean ages. The green
dotted line represents the median age, and the blue dashed line
represents the mean age, both located just above 45. This suggests the
sample is predominantly middle-aged, with a gradual decline in
respondent frequency after the age of 60. Few participants were over 80,
indicating a lower representation of older individuals

@@

\caption{\label{fig-one}Distribution of age given the 2018 survey
respondents. The mean and median of the sample is indicated with dashed
and dotted lines respectively.}

@@

\subparagraph{Comparison 2018 to 2021}\label{comparison-2018-to-2021}

Figure~\ref{fig-two} compares the age distributions between 2018 and
2021. The main difference is an increase in respondents aged 18-23 in
2021, with other age groups remaining relatively stable. This shift may
reflect changing demographic patterns or differences in survey
methodologies within the three years.

@@

\caption{\label{fig-two}This table highlights the age distribution of
survey respondents in 2018 and 2021.}

\paragraph{Education}\label{education}

\subparagraph{Individual 2018}\label{individual-2018-1}

Figure~\ref{fig-three} shows the distribution of respondents by their
highest level of education attained in 2018. The most common education
levels are ``Completed undergraduate degree'' and
``Postgraduate/professional school,'' indicating a well-educated
respondent population. In contrast, fewer respondents reported having
attended or completed community college, vocational, or trade school,
suggesting lower representation in these categories. A small portion of
respondents chose not to disclose their education level. Overall, the
distribution reflects a skew toward higher levels of educational
attainment.

@@

\subparagraph{Comparison 2018 to 2021}\label{comparison-2018-to-2021-1}

Figure~\ref{fig-four} compares the education levels of respondents in
2018 and 2021. The two surveys used different education categories,
making direct comparisons challenging. In 2018, 9\% of respondents had a
high school diploma or less, while in 2021, this figure rose to 38\%,
due to the addition of the ``less than high school'' and ``high school
or equivalent'' categories. This shift indicates that the 2021 survey
better represented individuals with lower educational attainment.

The 2018 survey provided more detailed categories, including options for
community college, trade school, and distinctions between ``some''
versus ``completed'' education. In contrast, the 2021 survey used
broader categories, which likely reduced the specificity of educational
data. Notably, the 2021 survey appears to have offered a more balanced
sample, with a wider spread across various education levels, while the
2018 sample was noticeably skewed toward respondents with higher
educational attainment. This suggests that the 2021 survey may have
captured a broader range of educational backgrounds, providing a more
well-rounded representation of Toronto's population.

@@

\caption{\label{fig-four}This table highlights the education
distribution of survey respondents in 2018 and 2021.''}

\paragraph{Extent Feel Informed}\label{extent-feel-informed}

\subparagraph{Individual 2018}\label{individual-2018-2}

Figure~\ref{fig-five} depicts the distribution of self-reported climate
change knowledge among survey respondents in 2018. The majority of
participants identified as ``Very informed,'' making this the most
prevalent category by a significant margin. A moderate number of
respondents reported being ``Not very informed,'' while a smaller yet
noticeable group described themselves as ``Extremely informed.'' In
contrast, the categories ``Not at all informed'' and ``Not very
informed'' have the lowest frequencies, with ``Not at all informed''
being particularly rare. This distribution suggests that most
respondents possess a relatively high level of awareness about climate
change, with very few indicating a lack of knowledge on the topic.

@@

\subparagraph{Comparison 2018 to 2021}\label{comparison-2018-to-2021-2}

Figure~\ref{fig-six} illustrates the self-reported level of extent
inviduals feel informed about climate change and climate action for 2018
and 2021. In 2018, the large marjority of individuals say they are very
informed with 60 percent and the follwoign category is not very informed
with 27 percent. In 2021, The majority of inividuals also say they are
very informed but to a lesser degree with 48 percent and this is closely
followed by only a little informed with 37 percent. There is a trend
that invidiusls in 2018 reported greater levels of knowldege about
climate change whereas 2021 illstrates a more even split between very
informed and only little informed, portraying that fewer individuals
have adequate knowledge about climate change.

@@

\caption{\label{fig-six}This table highlights the extent feel informed
distribution of survey respondents in 2018 and 2021.}

\paragraph{Best Method of Delivery}\label{best-method-of-delivery}

\subparagraph{Individual 2018}\label{individual-2018-3}

Figure~\ref{fig-seven}, shows 2018 survey results for the preferred
method for receiving climate change information. Advertising campaigns
emerged as the most popular choice, selected by approximately 55 percent
of respondents, followed closely by newsletters with 52 percent and the
Toronto.ca website with 46 percent. On the other hand, social media
platforms such as Facebook, Instagram, and Twitter were less popular
overall, with Twitter 12 percent and Instagram 17 percent garnering the
fewer votes. Notably, only 5 percent of respondents reported being
uninterested in receiving information about climate action. These
findings suggest that more traditional communication methods, such as
advertising and newsletters, resonated better with respondents compared
to social media.

@@

\subparagraph{Comparison 2018 to 2021}\label{comparison-2018-to-2021-3}

Figure~\ref{fig-eight} compares the preferred methods of delivering
climate change information between the 2018 and 2021 surveys. While the
top three methods namely advertising campaigns, newsletters, and the
Toronto.ca website remained consistent across both years, there were
notable shifts in preferences. Traditional methods such as newsletters,
advertising campaigns, brochures, and pamphlets saw a decline in 2021.
In contrast, social media platforms including Twitter, Facebook, and
Instagram gained traction, reflecting a growing reliance on digital
communication channels.

Additionally, the 2021 survey introduced new categories, including the
BetterHomesTO.ca website, which accounted for 15 percent of responses.
This addition underscores the increasing importance of specialized
online platforms in engaging the public on climate action. These changes
highlight a shift toward more diverse and digitally focused
communication strategies over time.

@@

\caption{\label{fig-eight}This table highlights the prefered method of
communication of survey respondents in 2018 and 2021.}

\paragraph{Likelihood Taking Action}\label{likelihood-taking-action}

Figure~\ref{fig-nine} illustrates the self-reported likelihood of
individuals taking various actions to minimize environmental impacts.
The x-axis displays nine different climate actions, while the y-axis
represents the percentage of individuals corresponding to each
likelihood level. A stacked bar chart is used to visualize the
distribution of responses, with a color-coded legend ranging from green
(high likelihood of action) to red (low likelihood of action).

Sorting waste into the correct bins is the most frequently reported
action that individuals are currently undertaking, accounting for 60
percent of responses. Other common actions individuals are already
participating in include walking or cycling short distances with 40
percent, reducing waste by contributing to the repurposing of used
products with 37 percent and reducing hydro usage with 36 percent

In contrast, the least popular actions include switching to electric or
hybrid vehicles, with 14 percent of respondents reporting they are very
unlikely to take this action and 27 percent stating they are somewhat
unlikely. Similarly, consuming less meat by incorporating more
plant-based foods is unpopular, with 12 percent reporting they are very
unlikely and 23 percent somewhat unlikely to adopt this change.
Minimizing car use is another action with low reported likelihood, with
10 percent of respondents very unlikely and 16 percent somewhat unlikely
to reduce their car usage.

@@

\subsection{Model}\label{sec-model}

The goal of our modeling strategy is twofold: first, to predict how
demographic factors influence individuals' likelihood of engaging in
climate-friendly behaviors, and second, to uncover the role of age and
education in shaping these actions. The actions considered include
investing in electric vehicles, reducing hydro usage, adopting meat
alternatives, and walking or cycling shorter distances. We use a
decision tree model to provide a clear, interpretable framework for
analyzing these relationships.

@@

Our model predicts the likelihood of engaging in specific
climate-positive actions, with age and education level serving as
predictors. Likelihood is measured on a five-point scale, ranging from
``Already doing this or have done this'' to ``Very unlikely.''

@@

We employ separate decision tree models for each action. Age
(continuous) and education (categorical) are the key predictors driving
predictions. Each tree splits the data at nodes based on these
variables, revealing demographic conditions associated with greater or
lesser likelihoods of climate-positive behaviors.
